## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}

```

We can also add else if statements in the middle if we need more options.

```{r}

```

Neat! We can also nest conditionals just like we do in Python.

```{r}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}

```

We can loop over more than just lists:

```{r}

```

Like Python, we have while loops in R:

```{r}

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}

```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
    #Question 1 Rock paper scissors
    rps <- function(p1, p2){
      if (p1 == p2) {
        return("Draw!")
      }
      
      # Determine the winner
      if ((p1 == "rock" && p2 == "scissors") ||
          (p1 == "scissors" && p2 == "paper") ||
          (p1 == "paper" && p2 == "rock")) {
        return("Player 1 won!")
      } else {
        return("Player 2 won!")
      }
    }

    #Question 2 Even or Odd
    even_or_odd <- function(n) {
       # Check if the number is even
      if (n %% 2 == 0) {
        return("Even")
      } else {
        return("Odd")
      }
    }

    #Question 3 Remove exclamation marks
    remove_exclamation_marks <- function(s){
      gsub("!", "", s)
    }

    #Question 4 convert number to string
    number_to_string <- function(n) {
     return(as.character(n))
    }

    #Question 5 sentence smash
    smash <- function(words) paste(words, collapse = " ")
    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
    fizzBuzzCheck <- function(input) {
      if (!is.numeric(input) || input%%1 != 0) {
        return("The input must be an integer.")
      }
      if (input %% 3 == 0 & input %% 5 == 0) {
        return("fizzbuzz")
      } else if (input %% 3 == 0) {
        return("fizz")
      } else if (input %% 5 == 0) {
        return("buzz")
      } else {
        return(input)
      }
    }

    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    marioStairs <- function(height) {
      if (!is.numeric(height) || height%%1 != 0) {
        cat("The input must be an integer.\n")
        return()
      }
      for (i in 1:height) {
        cat(strrep("X", i), "\n")
      }
    }
    marioStairs(4)

    ```
